Disk Heating, Galactoseismology, and the Formation of Stellar Halos

### 2017-06-27 paper/paper.tex
ally, ting comments

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{
$^{1}$ \quad Department of Astronomy, Columbia University \\
$^{2}$ \quad Department of Astrophysical Sciences, Princeton University \\
$^{3}$ \quad Max Planck Institute for Astronomy, Heidelberg\\
$^{4}$ \quad Department of Physics \& Astronomy, Texas A \& M University\\
$^{5}$ \quad City University of New York, LaGuardia Community College \\
$^{6}$ \quad Sydney Institute for Astronomy, School of Physics, University of Sydney \\
}

@@

This group sits in the north Galactic hemisphere in the area $\approx 125^\circ \lesssim l \lesssim 210^\circ$, $\approx 20^\circ \lesssim b \lesssim 40^\circ$ and at approximate distances of $\approx 12$--$20~\kpc$.

@@

distance dependence that is ignored in the quoted distance ranges above.

@@

% APW COMMENT: I found it strange that the historical context doesn't include
% our recent work - maybe we should state that this was the status before our
% group started working on follow-up, or something? Otherwise, GASS section
% should get a reference to upcoming Sheffield+(in prep.) paper on RR Lyrae,
% TriAnd section should get references to Sheffield+2014(?) and
% Price-Whelan+2015, A13 paper should get reference to Li+2017
% KVJ: hope sentence above helps?

@@

In our first results \cite{sheffield14}, we extended a prior sample of TriAnd
M giants \cite{rochapinto04} by obtaining spectra of all members identified
with an automated clustering algorithm applied to 2MASS catalog sources in the
region \cite{sharma10}.
We found that group members occupied two distinct loci in color-magnitude space, which we associated with the two MSTO identified in optical data by \cite{martin07}, named TriAnd 1 and TriAnd 2 respectively.
Stars in both TriAnd 1 and TriAnd 2  form  clear over-densities in radial velocity with low dispersion ($\sim 25~\kms$) compared to the background halo velocity distribution.
They both follow the same  sequence in velocity with a steady gradient of mean Galactic standard-of-rest (GSR) radial velocity ($v_{\rm GSR}$) from positive to negative $v_{\rm GSR}$ with increasing $l$.
% APW COMMENT: I added some more specifics, but it feels a little clunky.
% Trying to say that it goes from positive vGSR at low longitudes to negative
% vGSR at larger longitudes
%KVJ: slight re-write
In this work, we presented a model that simultaneously and approximately reproduces TriAnd 1 and 2 as tidal debris stripped over two separate pericentric passages from a single accreted satellite on a low-eccentricity, retrograde, near-planar orbit.
The debris structures in the simulations were morphologically closer to {\it streams} rather
than {\it shells}, but still subtended large areas on the sky as observed from
the Sun's position.

In subsequent work we continued the spectroscopic survey to look at A13, the new group M giant discovered through the group-finding algorithm \cite{sharma10}. A13 overlaps the TriAnd clouds in Galactic longitude, but is apparent in the Northern Galactic hemisphere at slightly brighter magnitudes. The spectra showed that, like the TriAnd clouds, this  group had a coherent velocity structure with low dispersion and a steady gradient with $l$, confirming the genuine association of its members \cite{li17}.

@@

\subsection{Stellar Populations}
 \label{sec:populations}
Motivated by the observed, low-dispersion velocity distribution of the M giant stars in TriAnd, we sought to observe other distance tracer stars --- RR Lyrae-type pulsators --- in the region, determine their membership, and improve the distance estimates to the structure.
We selected RR Lyrae stars in the TriAnd region from the pre-classified
variable star catalog produced by the Catalina Sky Survey \cite{drake14},
using a conservative distance cut to account for uncertainties in the RR Lyrae
and M giant photometric distance estimates.
We obtained spectra for $\approx$1/3 of the total number of RR Lyrae in the M
giant volume considered to be associated with TriAnd 1 and measured radial
velocities for these stars \cite{pricewhelan15}.

@@

Since we were unable to find any RR Lyrae clearly associated with TriAnd 1, our attempt to measure a more accurate distance to this structure was a failure.
However, the upper limit on the value of $\frrmg$ was in itself interesting.
RR Lyrae and M giants are tracers of populations with quite distinct metallicities: stars in the horizontal branch phase of evolution are typically only blue enough to cross the instability strip and become RR Lyrae if they have [Fe/H]$\lesssim -1.5$; and giant stars typically only evolve to colors red enough to become spectral class M if they have [Fe/H]$\gtrsim$-1.5.
Hence, a stellar population has to contain a significant range of metallicities in order to contain both types of stars.
%In order to aid with the interpretation of this discovery, Figure~\ref{fig:isochrones} plots isochrones for two $10~{\rm Gyr}$-old populations of low and intermediate metallicity, including the color range for RR Lyrae and M spectral class.
%It emphasizes our understanding that these two types of stars are tracers of populations with quite distinct metallicities and illustrates why, while the initial aim of our survey to find more accurate distances to TriAnd 1 and 2 had failed, our results could be even more intriguing.
The metallicity distributions in nearly all existing satellite galaxies
orbiting the Milky Way \cite[e.g.,][]{kirby11} are typically biased towards
low metallicities, such that they contain no or few M giant stars (i.e. $\frrmg = \infty$).
The largest satellites (the Large Magellanic Cloud and the Sagittarius dwarf galaxy) are exceptions as they contain the most metal rich populations; these dwarfs have $\frrmg \sim 0.5$ \cite{pricewhelan15}.
In contrast, the Galactic disk is an overall metal-rich population and thus has very few RR Lyrae \cite[i.e. $f_{\rm RR:M} \sim 0$][]{amrose01}, more consistent with our findings for the stellar population of TriAnd 1.

@@

\label{sec:abundances}
% APW COMMENT: This section needs more context and perhaps much less detail
% about the actual observations...

The origin of stellar associations can also be explored through the chemical compositions of their constituent stars.
However, our finding (described in Section \ref{sec:populations}) that the low-latitude-structures all have stellar populations (as indicated by $f_{\rm RR:M}$) that look more like the disk than known Galactic satellites is at odds with prior work on abundance patterns.
For example, high-resolution spectra of 21 M giants in Mon/GASS \cite{chou2010b} showed
[Ti/Fe] lower by up to 0.4 dex compared to the mean trends known for main-sequence stars of the Galactic disk (e.g., Reddy et al. 2003, Bensby et al. 2014) and a mean offset for [Y/Fe] of about 0.2 dex at ${\rm [Fe/H]} \approx -0.5$. A comparison with results from the same group for stars in the Sgr dwarf spheroidal galaxy \cite{chou2010a} suggested more similar patterns to those seen in Mon/GASS and support  an external origin for this structure. Subsequent work comes to similar conclusions for the TriAnd Clouds \cite{chou11}.
% APW COMMENT: there are hand-typed references in the paragraph above! We need to add to the bibliography

@@

In order to explore the apparent contradiction between the stellar populations and abundance work we have recently obtained high-resolution and signal-to-noise spectra of 15 stars in TriAnd 1 and A13 overdensities. 14 stars were observed with the HIRES-S spectrograph at the Keck-1 telescope \cite{vogt1994} and 1 star was observed using the the UVES spectrograph at the VLT (Program ID: 097.B-0770A).
%The spectral resolving power R of the HIRES spectra is 36\,000 and the UVES data have R $\sim$ 47\,000. All Keck spectra cover the full optical region, from $4800$ to $8770$ $\AA$, although the exact wavelength coverage varies as several slightly different instrument configurations were tried in an effort to get all the key lines into a single exposure. The signal-to-noise (SNR)/\AA of the HIRES spectra near 5200 \AA in the continuum at the center of the echelle order exceeds 200. For the UVES spectrum, the SNR of 50 near 5500 \AA was achieved. The fundamental parameters and chemical abundances of stars were determined using 2MASS and APASS photometry and the high-resolution spectra. All 15 stars are M-type giants with effective temperatures Teff ~ 3800 K and surface gravities log(g) ~ 1 dex. The $T_{\rm eff}$ estimates were derived using the Infrared Flux Method (Casagrande et al. 2010, 2014). Chemical abundances were measured for 6 chemical elements, including O, Na, Mg, Ca, Fe, Eu using LTE and the standard MARCS stellar model atmospheres (Gustafsson et al. 2008). We have also performed detailed analysis of the effects of non-local thermodynamic equilibrium (non-LTE) for the elements, where detailed calculations are available \cite{bergemann2011, bergemann2012, bergemann2016}, however, the NLTE corrections for the chosen lines are minor and do not change our conclusions. In what follows, we use our LTE results, because all studies of chemical abundances in dSph systems and most studies in the Milky Way to date have employed LTE with 1D hydrostatic models
We find that stars in the TriAnd 1 and in A13 overdensities have extremely similar chemical abundances, with the abundance dispersion $\leq 0.06$ dex for most chemical elements. The TriAnd 1 and A13 stars also have a very narrow metallicity spread,  ${\rm [Fe/H]} = -0.92 \pm 0.06$ dex. The abundances of all measured alpha-elements are uniformly enhanced at the level of 0.43 $\sim$ 0.06 dex for [O/Fe], 0.41 $\sim$ 0.02 dex for [Mg/Fe], $0.35 \pm 0.06$ dex for [Na/Fe], and 0.15 $\sim$ 0.03 dex for [Eu/Fe]. At this level, the TriAnd 1 and A13 stars are consistent with the abundances of the in-situ formed Milky Way thick disk stars \cite{fuhrmann2004,bergemann2014,bensby2014}.
\footnote{The only exception is [Ca/Fe], which is consistent with solar with the dispersion of 0.06 dex; however, the NLTE abundance correction for the \ion{Ca}{i} lines for M giants is 0.1 dex that will bring the results in agreement with the MW disk studies that are based on FGK dwarfs, for which the NLTE correction is much smaller \cite{Merle2011}}.
The abundance ratios are too high for the chemical abundance patterns observed in the stars of the Galactic satellites (Bonifacio et al. 2010, Shetrone et al. 2001, 2003, Tolstoy et al. 2009, de Boer et al. 2014), which are known to have low, typically solar ( [O/Fe], [Mg/Fe]) or even sub-solar ([Na/Fe]), ratios at [Fe/H] $\sim -1$.
% APW COMMENT: there are hand-typed references in the paragraph above! We need to add to the bibliography

@@

%Our results for TriAnd and A13 do not confirm the conclusions by \cite{chou2010b}. Although there could be several astrophysical reasons for the differences, this could be caused by the
One explanation for the disagreement between abundance work for stars in the same structure is
differences in the observed datasets and spectroscopic modeling techniques.
%Our observed data cover the full optical range from the near-UV to the near-IR, whereas
\cite{chou2010b} analysed a small wavelength region in the near-IR, limited to 150 \AA from 7440 to 7590 \AA. Because of this limitation, they could include only 11 \ion{Fe}{I} and 2 \ion{Ti}{I} lines in the determination of gravities, metallicities, and abundances. Our analysis, through a much wider wavelength coverage and high SNR attained for the observed spectra, allowed us to include 70 lines of \ion{Fe}{I} and \ion{Fe}{II} \cite{bergemann2012}, as well as 10 lines of \ion{Ti}{2}. Moreover,  \cite{bergemann2011} showed that \ion{Ti}{1} should be not be used in abundance studies because it is very sensitive to NLTE effects .
In order to explore the the sensitivity of abundance diagnostics to the line selection and wavelength regimes we performed test computations on our own data, using a reduced linelist. We have found that using the linelist from \cite{chou2010a}, the metallicities are over-estimated by $\sim 0.25$ dex and [Ti/Fe] are under-estimated by $0.4$ dex, compared to the results using the complete linelist. This suggests that the choice of the linelist and diagnostic spectral band (full optical or near-IR) as a plausible explanation  of the apparently discrepant results.
%{\it Summary: the abundance patterns  of the low-latitude structures are similar to those of the thick disk of our Galaxy.}

@@

% APW COMMENT: there are hand-typed references in the paragraph above! We need to add to the bibliography

@@

% APW COMMENT: If the sections are reorganized to be by structure instead, this
% should be moved to the section on Mon/GASS, I believe
While we have concentrated on follow-up studies of the known low-latitude structures as traced by M giants selected from 2MASS, knowledge of the spatial distribution of stars towards the anticenter region has been further refined using photometry in SDSS \cite{xu15} and PanSTARRS \cite{kaiser10,lurie17}.
Both these studies employed the novel technique of subtracting color-magnitude diagrams (CMD's) derived from fields in their photometric data which were symmetrically placed at equal and opposite Galactic latitudes and at the same Galactic longitudes.
These differenced CMD's allowed denser regions, closer to the Galactic plane and at smaller heliocentric distances  to be explored.
Both studies showed overdense arcs or stars oscillating between the northern and southern hemispheres as the distance from the Sun was increased towards the anticenter of our Galaxy.
The vast numbers of stars in these surveys allowed more clear identification of these global structures as clearly distinct, separate features.

@@

Coincident with these studies exploring structures at the very outer edge of our Galactic disk,
large scale spectroscopic surveys have allowed a detailed re-examination of the {\it local} distribution of stellar velocities.
Asymmetries between the Northern and Southern Galactic hemispheres in the vicinity of the Sun have been seen in the density and velocity distributions using data from SDSS \cite{widrow12,yanny13} and RAdial Velocity Experiment \cite[RAVE, see][]{??,williams13}.
Looking $\sim$2kpc out towards the Galactic anticenter, the Large Sky Area Multi-Object Fiber Spectroscopic Telescope \cite[LAMOST,][]{cui12,deng12,zhao12} finds similar asymmetries in radial and vertical velocities \cite{carlin13}.
The scale and sense of these asymmetries indicate moderate systematic motions (of order a few km/s)  of stars within the disk perpendicular to the plane,  suggesting both vertical movement of midplane, and compression and rarefaction of the vertical scale \cite[referred to as ``bending'' and ``breathing'' modes respectively --- see,][]{widrow14}.
It is natural to associate these asymmetries in motion as a local manifestations of the oscillations traced in space over much larger scales \cite{xu15,pricewhelan15}.
%{\it Summary: small-scale, systematic vertical motions of and within the disk have been detected in the Solar Neighborhood.}

@@

The observations summarized in Section \ref{sec:obs} indicate that:
\begin{itemize}
\item the low latitude structures --- Mon/GASS, TriAnd 1/2 and A13  --- each have low velocity dispersions supporting the genuine association of the stars within them;
\item Mon/GASS, TriAnd 1/2 and A13 also share a single, continuous sequence in average velocity as a function of $l$ suggestive of associations across these groups;
\item the stellar populations (as indicated by $f_{\rm RR:M}$) in these groups look like each other;
\item the stellar populations are more consistent with those in the Galactic disk rather than those observed in the stellar halo or Galactic  satellites;
\item the abundance patterns  of the low-latitude structures are similar to those of the thick disk of our Galaxy;
\item the low-latitude structures around the outer Galaxy can be connected to overdensities oscillating between the northern and southern hemispheres on smaller scales and traced all the way back to the Solar Neighborhood;
\item small-scale, systematic vertical motions of and within the disk have been detected in the Solar Neighborhood which suggest the local disk is moving up and down.
\end{itemize}

@@

Figure \ref{fig:cartoon} is a cartoon which summarizes all the observational works and their implications, by showing the approximate locations and amplitudes of the spatial and velocity structures that have been identified. (Note that the figure is misleading as the structures are {\it not} consistent morphologically with concentric rings that are axisymmetric about the Galactic center.)
Taken together, we conclude that: (i) there is compelling evidence that Mon/GASS, TriAnd 1/2 and A13 represent populations of stars formed in the disk that now exist at extreme $(R,Z)$;
(ii) these structures are associated with each other as part of a global system of vertical disk oscillations that can be traced all the way to the velocity asymmetries seen in the Solar Neighborhood; and
(iii) the stellar populations in these structures are inconsistent with a picture where they formed from an accreted satellite.

@@

Figure \ref{fig:chervin} illustrates these ideas with the results of simulations from our own recent work.
We extended the prior theoretical backdrop that looked at Mon/GASS to examine whether the extreme locations of TriAnd 1/2 could fit within the same picture using simulations of a disk disturbed (separately) by satellites on orbits that mimic those expected for the Large Magellanic Cloud and Sagittarius \cite{laporte17}.
With different masses and orbits (and consequently different interaction  strength, timings, and durartions) these satellite necessarily induced distinct but overlapping signatures.
In  more recent work, we found that a model that was capable of reproducing the scales of the observed disturbances (radial wavelength and  amplitude in space, as well as magnitude of offsets in velocity locally) needed:
the  interaction of Sgr with the disk of the Milky Way to be followed for several passages longer than prior work;
Sgr to have of sufficient initial mass and density to impact the disk in the last Gigayear with a remaining mass of $\sim3\times10^{9} M_{\odot}$;
and the disk to be realized with stars existing as far out as ???kpc form the Galactic center in order to populate the regions corresponding to TriAnd 1/2.
The interaction with the LMC modified the overall morphology of the structures induced, but was not sufficient alone to explain their properties.
The full details of these results will be discussed in an upcoming paper (Laporte et al, 2017).

@@

%It is interesting to place the current work in the context of ongoing and near-future studies of our Galaxy.
While the connections that have already been made between the different low-latitude structures and of these structures with the disk population are convincing, there are several directions for further observations which would greatly facilitate an informative comparison to theoretical work, allowing more detailed interpretations and tighter constraints on matter distributions and histories.

@@

\subsection{Other Galaxies}

@@

The growing samples of galaxies within and beyond the Local Group being mapped to extremely low surface brightness levels are intriguing.
Nearby, these can be reached, like the Milky Way, through star counts studies, most spectacularly for the case of our nearest neighbor, the Andromeda Galaxy, where giant star counts have revealed a significantly extended and richly featured outer stellar disk \cite{ferguson02,ibata05}.
Analogous studies have been carried out for galaxies up to distances of several Mpc \cite[e.g.,][]{monachesi13,crnojevi16}, although the focus of these studies has typically been on detecting the stellar halo of these objects.
Several dedicated surveys have made innovative advances in studying galaxies to LSB using a variety of techniques to reach limits below 30  mag/arcsec$^2$ in integrated light \citep[e.g.][]{delgado10,vandokkum14,duc15}.

>>>>

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{
$^{1}$ \quad Department of Astronomy, Columbia University, 550 W 120th st., Mail Code 5246, New York, NY, 10027 USA \\
$^{2}$ \quad Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08544, USA \\
$^{3}$ \quad Max Planck Institute for Astronomy, Heidelberg\\
$^{4}$ \quad Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA\\
$^{5}$ \quad Department of Natural Sciences, LaGuardia Community College, City University of New York, 31-10 Thomson Ave., Long Island City, NY 11101, USA \\
$^{6}$ \quad Sydney Institute for Astronomy, School of Physics, University of Sydney \\
}

@@

This group sits in the north Galactic hemisphere in the area $\approx 125^\circ
\lesssim l \lesssim 210^\circ$, $\approx 20^\circ \lesssim b \lesssim 40^\circ$
and at approximate distances of $\approx 12$--$20~\kpc$.
Follow-up spectroscopy again showed that the M giants in this structure have a
small velocity dispersion around a roughly linear trend with Galactic longitude
\cite{li17}

@@

distance that is ignored in the quoted distance ranges above.

@@

@@

In our first results \cite{sheffield14}, we extended a prior sample of TriAnd
M giants \cite{rochapinto04} by obtaining spectra of all candidate M giants identified by applying color-magnitude cuts to stars in the TriAnd region of the sky.
We found that the M giants occupied two distinct loci in color-magnitude space, which we associated with the two MSTO identified in optical data by \cite{martin07}, named TriAnd 1 and TriAnd 2 respectively.
Stars in both TriAnd 1 and TriAnd 2  form  clear over-densities in radial velocity with low dispersion ($\sim 25~\kms$) compared to the background halo velocity distribution.
They both follow the same  sequence in velocity with a steady gradient of mean Galactic standard-of-rest (GSR) radial velocity ($v_{\rm GSR}$) from positive to negative $v_{\rm GSR}$ with increasing $l$.
In this work, we presented a model that simultaneously and approximately reproduces TriAnd 1 and 2 as tidal debris stripped over two separate pericentric passages from a single accreted satellite on a low-eccentricity, retrograde, near-planar orbit.
The debris structures in the simulations were morphologically closer to {\it streams} rather
than {\it shells}, but still subtended large areas on the sky as observed from
the Sun's position.

In subsequent work we continued the spectroscopic survey to look at A13, the new group M giant discovered through the group-finding algorithm \cite{sharma10}. A13 overlaps the TriAnd clouds in Galactic longitude at one end and GASS at the other end, and is apparent in the Northern Galactic hemisphere at slightly brighter magnitudes than TriAnd. The spectra show that, like the TriAnd clouds, this group has a coherent velocity structure with low dispersion and a steady gradient with longitude, $l$, confirming the genuine association of its members \cite{li17}.

@@

\subsection{Stellar Populations}
 \label{sec:populations}
Motivated by the observed, low-dispersion velocity distribution of the M giant stars in TriAnd, we sought to observe other distance tracer stars --- RR Lyrae-type pulsators --- in the region, determine their membership, and improve the distance estimates to the structure.
We selected RR Lyrae stars in the TriAnd region from the Palomar Transient
Factory \cite[PTF][]{ptf}, using a conservative distance cut to account for
uncertainties in the RR Lyrae and M giant photometric distance estimates.
We obtained spectra for $\approx$1/3 of the total number of RR Lyrae in the M
giant volume considered to be associated with TriAnd 1 and measured radial
velocities for these stars \cite{pricewhelan15}.

@@

Since we were unable to find any RR Lyrae clearly associated with TriAnd 1, our attempt to measure a more accurate distance to this structure was unsuccessful.
However, the upper limit on the value of $\frrmg$ was in itself interesting.
RR Lyrae and M giants are tracers of populations with quite distinct metallicities: stars in the horizontal branch phase of evolution are typically only blue enough to cross the instability strip and become RR Lyrae if they have [Fe/H]$\lesssim -1.5$; and giant stars typically only evolve to colors red enough to become spectral class M if they have [Fe/H]$\gtrsim$-1.5.
Hence, a stellar population has to contain a significant range of metallicities in order to contain both types of stars.
%In order to aid with the interpretation of this discovery, Figure~\ref{fig:isochrones} plots isochrones for two $10~{\rm Gyr}$-old populations of low and intermediate metallicity, including the color range for RR Lyrae and M spectral class.
%It emphasizes our understanding that these two types of stars are tracers of populations with quite distinct metallicities and illustrates why, while the initial aim of our survey to find more accurate distances to TriAnd 1 and 2 had failed, our results could be even more intriguing.
The metallicity distributions in nearly all existing satellite galaxies
orbiting the Milky Way \cite[e.g.,][]{kirby11} are typically biased towards
low metallicities, such that they contain no or few M giant stars (i.e. $\frrmg = \infty$).
The largest satellites (the Large Magellanic Cloud and the Sagittarius dwarf galaxy) are exceptions as they contain the most metal rich populations; these dwarfs have $\frrmg \sim 0.5$ \cite{pricewhelan15}.
In contrast, the Galactic disk is an overall metal-rich population and thus has very few RR Lyrae \cite[i.e. $f_{\rm RR:M} \sim 0$][]{amrose01}, more consistent with our findings for the stellar population of TriAnd 1.

@@

\label{sec:abundances}

The origin of stellar associations can also be explored through the chemical compositions of their constituent stars.
However, our finding (described in Section \ref{sec:populations}) that the low-latitude-structures all have stellar populations (as indicated by $f_{\rm RR:M}$) that look more like the disk than known Galactic satellites is at odds with prior work on abundance patterns.
For example, high-resolution spectra of 21 M giants in Mon/GASS \cite{chou2010b} showed
[Ti/Fe] lower by up to 0.4 dex compared to the mean trends known for main-sequence stars of the Galactic disk \cite[e.g.,][]{reddy03,bensby2014} and a mean offset for [Y/Fe] of about 0.2 dex at ${\rm [Fe/H]} \approx -0.5$. A comparison with results from the same group for stars in the Sgr dwarf spheroidal galaxy \cite{chou2010a} suggested more similar patterns to those seen in Mon/GASS and support  an external origin for this structure. Subsequent work comes to similar conclusions for the TriAnd Clouds \cite{chou11}.

@@

% ALLY COMMENT: Was this the actual motivation, if we look back at the
% proposal? We were asking for time to explore the possibility of TriAnd being
% a kicked out disk population
In order to explore the apparent contradiction between the stellar populations and abundance work we have recently obtained high-resolution and signal-to-noise spectra of 15 stars in TriAnd 1 and A13 overdensities. 14 stars were observed with the HIRES-S spectrograph at the Keck-1 telescope \cite{vogt1994} and 1 star was observed using the the UVES spectrograph at the VLT (Program ID: 097.B-0770A).
%The spectral resolving power R of the HIRES spectra is 36\,000 and the UVES data have R $\sim$ 47\,000. All Keck spectra cover the full optical region, from $4800$ to $8770$ $\AA$, although the exact wavelength coverage varies as several slightly different instrument configurations were tried in an effort to get all the key lines into a single exposure. The signal-to-noise (SNR)/\AA of the HIRES spectra near 5200 \AA in the continuum at the center of the echelle order exceeds 200. For the UVES spectrum, the SNR of 50 near 5500 \AA was achieved. The fundamental parameters and chemical abundances of stars were determined using 2MASS and APASS photometry and the high-resolution spectra. All 15 stars are M-type giants with effective temperatures Teff ~ 3800 K and surface gravities log(g) ~ 1 dex. The $T_{\rm eff}$ estimates were derived using the Infrared Flux Method (Casagrande et al. 2010, 2014). Chemical abundances were measured for 6 chemical elements, including O, Na, Mg, Ca, Fe, Eu using LTE and the standard MARCS stellar model atmospheres (Gustafsson et al. 2008). We have also performed detailed analysis of the effects of non-local thermodynamic equilibrium (non-LTE) for the elements, where detailed calculations are available \cite{bergemann2011, bergemann2012, bergemann2016}, however, the NLTE corrections for the chosen lines are minor and do not change our conclusions. In what follows, we use our LTE results, because all studies of chemical abundances in dSph systems and most studies in the Milky Way to date have employed LTE with 1D hydrostatic models
We find that stars in the TriAnd 1 and in A13 overdensities have extremely similar chemical abundances, with the abundance dispersion $\leq 0.06$ dex for most chemical elements. The TriAnd 1 and A13 stars also have a very narrow metallicity spread,  ${\rm [Fe/H]} = -0.92 \pm 0.06$ dex. The abundances of all measured alpha-elements are uniformly enhanced at the level of 0.43 $\sim$ 0.06 dex for [O/Fe], 0.41 $\sim$ 0.02 dex for [Mg/Fe], $0.35 \pm 0.06$ dex for [Na/Fe], and 0.15 $\sim$ 0.03 dex for [Eu/Fe]. At this level, the TriAnd 1 and A13 stars are consistent with the abundances of the in-situ formed Milky Way thick disk stars \cite{fuhrmann2004,bergemann2014,bensby2014}.
\footnote{The only exception is [Ca/Fe], which is consistent with solar with the dispersion of 0.06 dex; however, the NLTE abundance correction for the \ion{Ca}{i} lines for M giants is 0.1 dex that will bring the results in agreement with the MW disk studies that are based on FGK dwarfs, for which the NLTE correction is much smaller \cite{Merle2011}}.
The abundance ratios are too high for the chemical abundance patterns observed in the stars of the Galactic satellites \cite{Bonifacio:2000,shetrone2001,shetrone2003,Tolstoy:2009,deBoer:2014}, which are known to have low, typically solar ( [O/Fe], [Mg/Fe]) or even sub-solar ([Na/Fe]), ratios at [Fe/H] $\sim -1$.

@@

%Our results for TriAnd and A13 do not confirm the conclusions by \cite{chou2010b}. Although there could be several astrophysical reasons for the differences, this could be caused by the
One explanation for the disagreement between abundance work for stars in the same structure is
differences in the observed datasets and spectroscopic modeling techniques.
%Our observed data cover the full optical range from the near-UV to the near-IR, whereas
\cite{chou2010b} analysed a small wavelength region in the near-IR, limited to 150 \AA\ from 7440 to 7590 \AA. Because of this limitation, they could include only 11 \ion{Fe}{I} and 2 \ion{Ti}{I} lines in the determination of gravities, metallicities, and abundances. Our analysis, through a much wider wavelength coverage and high SNR attained for the observed spectra, allowed us to include 70 lines of \ion{Fe}{I} and \ion{Fe}{II} \cite{bergemann2012}, as well as 10 lines of \ion{Ti}{2}. Moreover,  \cite{bergemann2011} showed that \ion{Ti}{1} should be not be used in abundance studies because it is very sensitive to NLTE effects .
In order to explore the sensitivity of abundance diagnostics to the line selection and wavelength regimes we performed test computations on our own data, using a reduced linelist. We have found that using the linelist from \cite{chou2010a}, the metallicities are over-estimated by $\sim 0.25$ dex and [Ti/Fe] are under-estimated by $0.4$ dex, compared to the results using the complete linelist. This suggests that the choice of the linelist and diagnostic spectral band (full optical or near-IR) as a plausible explanation  of the apparently discrepant results.
%{\it Summary: the abundance patterns  of the low-latitude structures are similar to those of the thick disk of our Galaxy.}

@@

@@

% APW COMMENT: If the sections are reorganized to be by structure instead, this
% should be moved to the section on Mon/GASS, I believe
While we have concentrated on follow-up studies of the known low-latitude structures as traced by M giants selected from 2MASS, knowledge of the spatial distribution of MSTO stars towards the anticenter region has been further refined using photometry from the SDSS \cite{xu15} and PS1 surveys (Lurie et al., in prep.).
Both of these studies employ the novel technique of subtracting color-magnitude diagrams (CMD's) derived from fields in their photometric data which were symmetrically placed at equal and opposite Galactic latitudes and at the same Galactic longitudes.
These differenced CMD's allowed denser regions, closer to the Galactic plane and at smaller heliocentric distances  to be explored.
Both studies showed overdense arcs or stars oscillating between the northern and southern hemispheres as the distance from the Sun was increased towards the anticenter of our Galaxy.
The vast numbers of stars in these surveys allowed more clear identification of these global structures as clearly distinct, separate features.

@@

Coincident with these studies exploring structures at the very outer edge of our Galactic disk,
large scale spectroscopic surveys have allowed a detailed re-examination of the {\it local} distribution of stellar velocities.
Asymmetries between the Northern and Southern Galactic hemispheres in the vicinity of the Sun have been seen in the density and velocity distributions using data from SDSS \cite{widrow12,yanny13} and RAdial Velocity Experiment \cite[RAVE, see][]{??,williams13}.
Looking $\sim$2 kpc out towards the Galactic anticenter, the Large Sky Area Multi-Object Fiber Spectroscopic Telescope \cite[LAMOST,][]{cui12,deng12,zhao12} finds similar asymmetries in radial and vertical velocities \cite{carlin13}.
The scale and sense of these asymmetries indicate moderate systematic motions (of order a few km/s)  of stars within the disk perpendicular to the plane,  suggesting both vertical movement of the midplane, and compression and rarefaction of the vertical scale \cite[referred to as ``bending'' and ``breathing'' modes respectively --- see,][]{widrow14}.
It is natural to associate these asymmetries in motion as a local manifestations of the oscillations traced in space over much larger scales \cite{xu15,pricewhelan15}.
%{\it Summary: small-scale, systematic vertical motions of and within the disk have been detected in the Solar Neighborhood.}

@@

The observations summarized in Section \ref{sec:obs} indicate that:
\begin{itemize}
\item the low latitude structures --- Mon/GASS, TriAnd 1/2 and A13  --- each have low velocity dispersions supporting the genuine association of the stars within them;
\item Mon/GASS, TriAnd 1/2 and A13 also share a single, continuous sequence in average velocity as a function of $l$ suggestive of associations across these groups;
\item the stellar populations (as indicated by $f_{\rm RR:M}$) in these groups look like each other;
% ALLY COMMENT: Combine with next point? “look like each other” is a bit vague!
\item the stellar populations are more consistent with those in the Galactic disk rather than those observed in the stellar halo or Galactic  satellites;
\item the abundance patterns  of the low-latitude structures are similar to those of the thick disk of our Galaxy;
\item the low-latitude structures around the outer Galaxy can be connected to overdensities oscillating between the northern and southern hemispheres on smaller scales and traced all the way back to the Solar Neighborhood;
\item small-scale, systematic vertical motions of and within the disk have been detected in the Solar Neighborhood which suggest the local disk is moving up and down.
\end{itemize}

@@

% TING COMMENT: everything above line 251 is really summary of observation.
% Maybe this should go to Section 2.6? It is kinda confusing that this section
% is for theory but it were all about observations for the first 20 lines
Figure \ref{fig:cartoon} is a cartoon which summarizes all the observational works and their implications, by showing the approximate locations and amplitudes of the spatial and velocity structures that have been identified. (Note that the figure is misleading as the structures are {\it not} consistent morphologically with concentric rings that are axisymmetric about the Galactic center.)
Taken together, we conclude that: (i) there is compelling evidence that Mon/GASS, TriAnd 1/2 and A13 represent populations of stars formed in the disk that now exist at extreme $(R,Z)$;
(ii) these structures are associated with each other as part of a global system of vertical disk oscillations that can be traced all the way to the velocity asymmetries seen in the Solar Neighborhood; and
(iii) the stellar populations in and chemical abundances of these structures are inconsistent with a picture in which they formed from an accreted satellite.

@@

Figure \ref{fig:chervin} illustrates these ideas with the results of simulations from our own recent work.
We extended the prior theoretical backdrop that looked at Mon/GASS to examine whether the extreme locations of TriAnd 1/2 could fit within the same picture using simulations of a disk disturbed (separately) by satellites on orbits that mimic those expected for the Large Magellanic Cloud and Sagittarius \cite{laporte17}.
With different masses and orbits (and consequently different interaction  strength, timings, and durartions) these satellite necessarily induced distinct but overlapping signatures.
In  more recent work, we found that a model that was capable of reproducing the scales of the observed disturbances (radial wavelength and  amplitude in space, as well as magnitude of offsets in velocity locally) needed:
the  interaction of Sgr with the disk of the Milky Way to be followed for several passages longer than prior work;
Sgr needs to have sufficient initial mass and density to impact the disk in the last Gigayear with a remaining mass of $\sim3\times10^{9} M_{\odot}$;
and the disk to be realized with stars existing as far out as ???kpc form the Galactic center in order to populate the regions corresponding to TriAnd 1/2.
The interaction with the LMC modified the overall morphology of the structures induced, but was not sufficient alone to explain their properties.
The full details of these results will be discussed in an upcoming paper (Laporte et al, 2017).

@@

%It is interesting to place the current work in the context of ongoing and near-future studies of our Galaxy.
While the connections that have already been made between the different low-latitude structures and the disk population are convincing, there are several directions for further observations which would greatly facilitate an informative comparison to theoretical work, allowing more detailed interpretations and tighter constraints on matter distributions and histories.

@@

% TING COMMENT: There are also a significant amount of studies on "galaxy
% morphology and warps" with distance galaxies. For example:
% http://adsabs.harvard.edu/abs/2016JKAS...49..239A
% You may want to mention these studies (reference therein) as well....
\subsection{Other Galaxies}

@@

The growing samples of galaxies within and beyond the Local Group being mapped to extremely low surface brightness levels are intriguing.
Nearby, these can be reached, like the Milky Way, through star counts studies, most spectacularly for the case of our nearest neighbor, the Andromeda Galaxy, where giant star counts have revealed a significantly extended and richly featured outer stellar disk \cite{ferguson02,ibata05}.
Analogous studies have been carried out for galaxies up to distances of several Mpc \cite[e.g.,][]{monachesi13,crnojevi16}, although the focus of these studies has typically been on detecting the stellar halo of these objects.
Several dedicated surveys have made innovative advances in studying galaxies to low surface-brightness using a variety of techniques to reach limits below 30 mag/arcsec$^2$ in integrated light \citep[e.g.][]{delgado10,vandokkum14,duc15}.
